### 2019-04-05 Thesis_Latex/ThesisFinal/Chapter2/Chapter2.tex
update final thesis

\chapter{Literature Review}
In this chapter we introduce the literature survey of the various algorithms used for robot navigation. The navigation methods are divided into two kinds of control: global path planning and local motion control. First we analyze these two types of strategies and then we describe the current situation of Model Prediction which is the control method that we have adopted.
\section{Global Path Searching Method}
Global path planning uses information about a priori model or a map of the enviroment to evaluate the shortest path that allows the motion from a starting point to the goal position.
There are a lot of path planning algorithms, like cell decomposition, road map or potential field, where the calculation of a complete trajectory from a start position to one or more goal positions can be computed off-line. However they produce a reliable path if and only if the map of the enviroment is already available. So, this prior knowledge of the space where the robot can move, must be accessible.
The most famous method developed for global navigation is the Dijkstra algorithm \cite{Dijkstra:1959}. In recent years, an evolution of this method called A* is one of the most used \cite{Hart:1968}: this algorithm gives a complete and optimal global path in static enviroments. However, it was upgraded in D* \cite{Stentz93optimaland} for efficient online searching of a dynamic enviroment, which gives sequences of path points in the known or partially known space. The basic idea of these two methods is the following: minimizing their cost functions, these strategies have the capability of a quickly  redesign when the conditions of the space change, to guarantee an optimal solution of the trajectory from the starting point to the goal location. Moreover it has been shown that the D* algorithm is much more efficent than the A* method. Now we will analyze the general structure of these two algorithms and the practical operation based on a representation of an environment grid around the vehicle. However these methods are not sufficient to create a complete robot system navigation. We will see later that if the information of the map is not yet available, the system has to use onboard sensors to define a path and avoid obstacles.

@@

The most famous directional approaches (generate a direction for the robot/vehicle) are:
\begin{itemize}
	\item Potential field method \cite{Khatib1986}, where the robot is represented as a particle and it is subject to forces that are produced by the surrounding environment;
	\item Virtual Force Field which expands to Vector Field Histogram \cite{Borenstein1991}, where it utilizes a statistical representation of the vehicle's space through the so-called histogram grid;
	\item Nearness Diagram algorithm \cite{Montano2000}, which  performs a high level information extraction and interpretation of the environment, used to generate the motion commands;
\end{itemize}
while the velocity space approaches, that manage the robot/vehicle considering translation and rotation velocities, are:
\begin{itemize}
	\item Curvature Velocity method \cite{Simmons1996} which formulates the problem as one of constrained optimization in velocity space;
	\item Dynamic Window method \cite{Fox1997} which is divided in two main phases; in first one, it generates a valid search space while in the second one it selects an optimal solution in the search space.
\end{itemize}
Now we will analyze the general structure of some of these algorithms. In particular it is important to underline that to create a complete navigation system it is necessary that there are both local algorithms and global methods such as those seen previously.

\subsection{Potential Field Method}
The first local method that we analyzed is the Potential Field; in this strategy the robot/vehicle is represented as a particle that moves in the workspace and it is affected by attractive and repulsive forces that are generated by the surrounding environment. In particular the target propagates an attractive force while the obstacles transmit a repulsive forces. The latter can be evaluated considering only the distance from the obstacles and the instantaneous vehicle velocity and accelerations. Combining these two forces it is possible to compute the trajectory of the robot at every time instant. The total force can be used as input to control the trajectory of the vehicle.
\subsection{Curvature Velocity Method}
Another obstacle avoidance problem can be solved applying the
Curvature Velocity Method where the problem becomes a constrained optimization in the velocity space of a syncro-drive steered robot. This strategy tries to maximize
the following function:
\begin{equation}
\label{eqn:CVM}
	f(tv,rv)=\alpha_1\text{speed}(tv)+\alpha_2\text{dist}(tv,rv)+\alpha_3\text{head}(rv)
\end{equation}
where $tv$ and $rv$ are respectively the robot’s translational and rotational velocities while $\text{speed}(tv)$ represents the speed of the vehicle, $\text{head}(rv)$ is the heading of the robot and $\text{dist}(tv,rv)$ is distance to a set of obstacles within a given limit. These last three terms can be expressed as follows:
\begin{equation}
	\text{speed}(tv) = \frac{tv}{tv_{\text{max}}}\qquad 
	\text{head}(tv) = 1-\frac{\theta_g-rvT_c}{\pi}\qquad
	\text{dist}(tv,rv)=\frac{D_{\text{limit}}(tv,rv,OBS)}{\text{limit}}.
\end{equation}
At each sampling time $T_c$, the constrained optimization problem is solved choosing the transational and rotational velocities to adjuste $\alpha_1,\alpha_2$ and $\alpha_3$ values. In order to limit the search space of possible curvatures of the vehicle in velocity space, the value $D_\text{limit}$ is bounded and discretized. Finally we can classify these intervals in disjoint, contained by, contains or overlapping and they are fundamentals to compute optimal values for the distance function and consequently for the equation (\ref{eqn:CVM}).

@@

\section{MPC in Autonomous Driving}
The algorithms we have developed in this project have a different approach than those mentioned above. In this section we have reported some algorithms based on the MPC used in the field of autonomous navigation. This literature review was fundamental to understand what methods exist and the strategies adopted to solve certain problems.
In \cite{Frasch2013} the authors addressed the problem of real-time obstacle avoidance on low-friction road surfaces using spatial Nonlinear Model Predictive Control (NMPC). They considered a nonlinear four wheel vehicle dynamics model that includes load transfer and they also used the ACADO Code Generation tool to generate NMPC algorithms based on the real-time iteration scheme for dynamic optimization. An obstacle avoidance problem solved with an MPC approach that integrates the artificial potential field method is discussed in \cite{Manzoor2015} where the controller is combined with the feedback linearization in order to manage the control problem of a single robot with unicycle kinematics and collision avoidance function. However in many traffic emergency situations a collision cannot be prevented by braking alone. This is the reason why in \cite{Werling2012} the authors proposed an obstacle avoidance method based on the NMPC paradigm that simultaneously optimizes steering and braking. In \cite{Multi2017} a path planning and tracking framework is considered to maintain a collision-free path for autonomous vehicles. In particular to track the planned
trajectory, the proposed controller formulated the tracking task as a multiconstrained model predictive control (MMPC) problem and evaluated the steering angle to prevent the robot  from colliding with a moving obstacle vehicle. Finally we considered another driver assistance method for obstacle avoidance based on constrained model predictive control. The algorithm also in this case was reduced to a convex quadratic programming problem. \cite{Wada2017}. All these methods, based on the MPC paradigm, allowed us to define two different situations where the use this type of controller is required. The first situation we have faced is a collision avoidance problem considering moving obstacles. In the second scenario we have instead built a system that allows tracking of the lane. These two approaches, faced earlier, have been solved through a new type of MPC called Adaptive MPC.

>>>>

\chapter{Literature Review}
In this chapter we introduce the literature review of the various algorithms used for robot navigation. The navigation methods are divided into two kinds of control: global path planning and local motion control. First we analyze these two types of strategies and then we describe the current situation of Model Prediction which is the control method that we have adopted.
\section{Global Path Searching Method}
Global path planning uses information about a priori model or a map of the enviroment to evaluate the shortest path that allows the motion from a starting point to the goal position.
There are a lot of path planning algorithms, like cell decomposition, road map or potential field, where the calculation of a complete trajectory from a start position to one or more goal positions can be computed off-line. However they produce a reliable path if and only if the map of the enviroment is already available. So, this prior knowledge of the space where the robot can move, must be accessible.
The most famous method developed for global navigation is the Dijkstra algorithm \cite{Dijkstra:1959}. In recent years, an evolution of this method called A* is one of the most used \cite{Hart:1968}: this algorithm gives a complete and optimal global path in static enviroments. However, it was upgraded in D* \cite{Stentz93optimaland} for efficient online searching of a dynamic enviroment, which gives sequences of path points in the known or partially known space. The basic idea of these two methods is the following: minimizing their cost functions, these strategies have the capability of a quickly  redesign when the conditions of the space change, to guarantee an optimal solution of the trajectory from the starting point to the goal location. Moreover it has been shown that the D* algorithm is much more efficent than the A* method. Now we will analyze the general structure of these two algorithms and the practical operation based on a representation of an environment grid around the vehicle. However these methods are not sufficient to create a complete robot system navigation. We will see later that if the information of the map is not yet available, the system has to use onboard sensors to define a path and avoid obstacles.

@@

The most famous directional approaches (generate a direction for the robot/vehicle) are:
\begin{itemize}
	\item Potential field method \cite{Khatib1986}, where the robot is represented as a particle and it is subject to forces that are produced by the surrounding environment;
	\item Virtual Force Field which expands to Vector Field Histogram \cite{Borenstein1991}, where it utilizes a statistical representation of the vehicle's space through the so-called histogram grid;
	\item Nearness Diagram algorithm \cite{Montano2000}, which  performs a high level information extraction and interpretation of the environment, used to generate the motion commands;
\end{itemize}
while velocity space approaches, that manage the robot/vehicle considering translation and rotation velocities, are:
\begin{itemize}
	\item Curvature Velocity method \cite{Simmons1996} which formulates the problem as one of constrained optimization in velocity space;
	\item Dynamic Window method \cite{Fox1997} which is divided in two main phases; in first one, it generates a valid search space while in the second one it selects an optimal solution in the search space.
\end{itemize}
Now we will analyze the general structure of some of these algorithms. In particular it is important to underline that to create a complete navigation system it is necessary that there are both local algorithms and global methods such as those seen previously.

\subsection{Potential Field Method}
The first local method that we analyzed is the Potential Field; in this strategy the robot/vehicle is represented as a particle that moves in the workspace and it is affected by attractive and repulsive forces that are generated by the surrounding environment. In particular the target propagates an attractive force while the obstacles transmit a repulsive forces. The latter can be evaluated considering only the distance from the obstacles and the instantaneous vehicle velocity and accelerations. Combining these two forces it is possible to compute the trajectory of the robot at every time instant. The total force can be used as input to control the trajectory of the vehicle.
\subsection{Curvature Velocity Method}
Another obstacle avoidance problem can be solved applying the
Curvature Velocity Method where the problem becomes a constrained optimization in the velocity space of a syncro-drive steered robot \cite{CVM2003}. This strategy tries to maximize
the following function:
\begin{equation}
\label{eqn:CVM}
	f(tv,rv)=\alpha_1\text{speed}(tv)+\alpha_2\text{dist}(tv,rv)+\alpha_3\text{head}(rv)
\end{equation}
where $tv$ and $rv$ are respectively the robot’s translational and rotational velocities while $\text{speed}(tv)$ represents the speed of the vehicle, $\text{head}(rv)$ is the heading of the robot and $\text{dist}(tv,rv)$ is distance to a set of obstacles within a given limit. These last three terms can be expressed as follows:
\begin{equation}
	\text{speed}(tv) = \frac{tv}{tv_{\text{max}}}\qquad 
	\text{head}(tv) = 1-\frac{\theta_g-rvT_c}{\pi}\qquad
	\text{dist}(tv,rv)=\frac{D_{\text{limit}}(tv,rv,OBS)}{\text{limit}}.
\end{equation}
At each sampling time $T_c$, the constrained optimization problem is solved choosing the transational and rotational velocities to adjuste $\alpha_1,\alpha_2$ and $\alpha_3$ values. In order to limit the search space of possible curvatures of the vehicle in velocity space, the value $D_\text{limit}$ is bounded and discretized. Finally we can classify these intervals in disjoint, contained by, contains or overlapping and they are fundamentals to compute optimal values for the distance function and consequently for the equation (\ref{eqn:CVM}).

@@

\section{MPC in Autonomous Driving}
The algorithms we have developed in this project have a different approach than those mentioned above. In this section we have reported some algorithms based on the MPC used in the field of autonomous navigation. This literature review was fundamental to understand what methods exist and the strategies adopted to solve certain problems.
In \cite{Frasch2013} the authors addressed the problem of real-time obstacle avoidance on low-friction road surfaces using spatial Nonlinear Model Predictive Control (NMPC). They considered a nonlinear four wheel vehicle dynamics model that includes load transfer and they also used the ACADO Code Generation tool to generate NMPC algorithms based on the real-time iteration scheme for dynamic optimization. An obstacle avoidance problem solved with an MPC approach that integrates the artificial potential field method is discussed in \cite{Manzoor2015} where the controller is combined with the feedback linearization in order to manage the control problem of a single robot with unicycle kinematics and collision avoidance function. However in many traffic emergency situations a collision cannot be prevented by braking alone. This is the reason why in \cite{Werling2012} the authors proposed an obstacle avoidance method based on the NMPC paradigm that simultaneously optimizes steering and braking. In \cite{Multi2017} a path planning and tracking framework is considered to maintain a collision-free path for autonomous vehicles. In particular to track the planned
trajectory, the proposed controller formulated the tracking task as a multiconstrained model predictive control (MMPC) problem and evaluated the steering angle to prevent the robot  from colliding with a moving obstacle vehicle. Finally we considered another driver assistance method for obstacle avoidance based on constrained model predictive control. The algorithm also in this case was reduced to a convex quadratic programming problem. \cite{Wada2017}. All these methods, based on the MPC paradigm, allowed us to define two different situations where the use of this type of controller is required. The first situation we have faced is a collision avoidance problem considering moving obstacles. In the second scenario we have instead built a system that allows tracking of the lane. These two approaches, faced earlier, have been solved through a new type of MPC called Adaptive MPC.

### 2019-04-05 Thesis_Latex/ThesisFinal/Chapter3/Chapter3.tex
update final thesis

\section{Generic Model Predictive Control problem}
Model Predictive Control (MPC), also known as Moving Horizon Control (MHC) or Receding Horizon Control (RHC), is a popular method for the control of slow dynamical systems, to generate the required
control inputs that are calculated at each sampling instance $k$, using the current state as initial
conditions to solve a finite optimal control problem. Some of the advantages of using MPC are:
\begin{itemize}
\item the ability to handle unstable, time variable, non-minimum phase systems;
\item robustness feature with the uncertainties in the nonlinear systems;
\item built in feed-forward control to handle disturbances in the processes;
\item enhanced tuning features to achieve the best response including transient responses;
\item the possibility to introduce constraints in a natural form;
\item if the references are known in advance, they can be used in order to optimize the reference tracking.
\end{itemize}

The methodology of all the controllers belonging to the MPC family is characterized by the following strategy, represented in Figure \ref{fig:mpc_theory}. The future outputs for a  determined horizon, called  the  prediction horizon, are predicted at each instant $k$ using the process model. These predicted outputs depend on the known values up to instant $k$ (past inputs and outputs) and on the future control signals which are those to be sent to the system and calculated. The set of future control signals is calculated by optimizing a determined criterion to keep the process as close as possible to the reference trajectory. This criterion usually takes the form of a quadratic function of the errors between the predicted output signal and the predicted reference trajectory. The control effort is included in the objective function in most
cases. 
\begin{figure}[!h]
	\centering
	\includegraphics[width=\textwidth]{./figure/mpc_theory.png}
	\caption{A discrete Model Predictive Control scheme adapted from \cite{mpctoolbox}.}
	\label{fig:mpc_theory}
\end{figure}

@@

MPC is typically formulated in the state space form. For a given discrete linear time-invariant (LTI) system:
\begin{equation}
\label{eqn:MPC_plant_discrete}
\vec{x}(k+1)=\vec{A}\vec{x}(k)+ \vec{B} \vec{u}(k)
\end{equation}
where $\vec{x}(k)\in\mathbb{R}^n$, $\vec{u}(k)\in\mathbb{R}^m$ are the state and the input, respectively. The central idea in the Model Predictive Control is to minimize some cost function, while still ensuring that some constraints are fulfilled. The generic MPC problem can be written as follows:
\begin{equation}
\label{eqn:MPC_optimization}
\begin{aligned}
& \underset{\textbf{u}}{\text{minimize}}
& & J(\vec{x}(k), \textbf{u}) \\
& \text{subject to}
& & \vec{x}_{k+i+1} = \vec{A}\vec{x}_{k+i}+ \vec{B} \vec{u}_{k+i}\quad\forall i=0,\dots N-1;\\
& & & \vec{x}_{k+i}\in \mathbb{X}\quad\forall i=0,\dots N-1;\\
& & & \vec{u}_{k+i}\in \mathbb{U}\quad\forall i=0,\dots N-1;\\  
& & & \vec{x}_{k+N}\in \mathbb{X}_f;\quad\vec{x}_k = \vec{x}(k).
\end{aligned}
\end{equation}
where $\textbf{u}=(\vec{u}_k,\dots,\vec{u}_{k+N-1})$ is a sequence of control inputs, $\vec{x}_{k+i}$ is the state at time $k+i$ as predicted at time $k$, and $N$ is the prediction horizon. The sets $\mathbb{X}\in\mathbb{R}^n$ and $\mathbb{U}\in\mathbb{R}^m$ define the constraints on the state and the input, respectively. Finally, the set $\mathbb{X}_f\subseteq\mathbb{X}$ defines the terminal constraint on the state. If we consider a regulation problem, the system (\ref{eqn:MPC_plant_discrete}) should be steered to the origin and the cost function $J(\vec{x}(k), \textbf{u})$ could be in a quadratic form as follows:
\begin{equation}
\label{eqn:MPC_cost_function_regulation}
	J(\vec{x}(k), \textbf{u}) = \vec{x}^\intercal_{k+N}\vec{P}_f\vec{x}_{k+N}+\sum_{i=1}^{N}\Big(\vec{x}^\intercal_{k+i}\vec{Q}\vec{x}_{k+i}+\vec{u}_{k+i}^\intercal\vec{R}\vec{u}_{k+i}\Big)
\end{equation}
where $\vec{P}_f,\vec{Q}\geq0$ (positive semi-definite) and $\vec{R}>0$ (positive definite) are weighting matrices.

@@

An MPC has many strengths. Given that the model is discrete and linear it handles multivariable problems very well. Also mathematical convexity is an important part of the resulting problem formulation of an MPC. In fact there exists efficient solvers for convex optimization problems but it is therefore desirable that the MPC problem (\ref{eqn:MPC_optimization}) is convex which is ensured if:
\begin{enumerate}
\item the cost function is convex;
\item the prediction model is linear;
\item the constraint sets $\mathbb{X}, \mathbb{U}$ are convex.	
\end{enumerate}	
The optimization handles actuator constraints and state constraints naturally in the optimization which allows for
the process to be operated much closer to the hard constraints, which improves control performance and efficiency. Because of its predictive nature it is able to solve a variety of problems and handle disturbances smoothly.

>>>>

\section{Generic Model Predictive Control problem}
Model Predictive Control (MPC), also known as Moving Horizon Control (MHC) or Receding Horizon Control (RHC), is a popular method for the control of slow dynamical systems, to generate the required
control inputs that are calculated at each sampling instance $k$, using the current state as initial
conditions to solve a finite optimal control problem.\\
Some of the advantages of using MPC, reported in \cite{camacho2004}, are:
\begin{itemize}
\item the ability to handle unstable, time variable, non-minimum phase systems;
\item robustness feature with the uncertainties in the nonlinear systems;
\item built in feed-forward control to handle disturbances in the processes;
\item enhanced tuning features to achieve the best response including transient responses;
\item the possibility to introduce constraints in a natural form;
\item if the references are known in advance, they can be used in order to optimize the reference tracking.
\end{itemize}

The methodology of all the controllers belonging to the MPC family is characterized by the following strategy, represented in Figure \ref{fig:mpc_theory}. The future outputs for a  determined horizon, called  the  prediction horizon, are predicted at each instant $k$ using the process model. These predicted outputs depend on the known values up to instant $k$ (past inputs and outputs) and on the future control signals which are those to be sent to the system and calculated. The set of future control signals is calculated by optimizing a determined criterion to keep the process as close as possible to the reference trajectory. This criterion usually takes the form of a quadratic function of the errors between the predicted output signal and the predicted reference trajectory. The control effort is included in the objective function in most cases \cite{camacho2013model}. 
\begin{figure}[!h]
	\centering
	\includegraphics[width=\textwidth]{./figure/mpc_theory.png}
	\caption{A discrete Model Predictive Control scheme adapted from \cite{mpctoolbox}.}
	\label{fig:mpc_theory}
\end{figure}

@@

MPC is typically formulated in the state space form. For a given discrete linear time-invariant (LTI) system:
\begin{equation}
\label{eqn:MPC_plant_discrete}
\vec{x}(k+1)=\vec{A}\vec{x}(k)+ \vec{B} \vec{u}(k)
\end{equation}
where $\vec{x}(k)\in\mathbb{R}^n$, $\vec{u}(k)\in\mathbb{R}^m$ are the state and the input, respectively. The central idea in the Model Predictive Control is to minimize some cost function, while still ensuring that some constraints are fulfilled. The generic MPC problem can be written as follows:
\begin{equation}
\label{eqn:MPC_optimization}
\begin{aligned}
& \underset{\textbf{u}}{\text{minimize}}
& & J(\vec{x}(k), \textbf{u}) \\
& \text{subject to}
& & \vec{x}_{k+i+1} = \vec{A}\vec{x}_{k+i}+ \vec{B} \vec{u}_{k+i}\quad\forall i=0,\dots N-1;\\
& & & \vec{x}_{k+i}\in \mathcal{X}\quad\forall i=0,\dots N-1;\\
& & & \vec{u}_{k+i}\in \mathcal{U}\quad\forall i=0,\dots N-1;\\  
& & & \vec{x}_{k+N}\in \mathcal{X}_f;\quad\vec{x}_k = \vec{x}(k).
\end{aligned}
\end{equation}
where $\textbf{u}=(\vec{u}_k,\dots,\vec{u}_{k+N-1})$ is a sequence of control inputs, $\vec{x}_{k+i}$ is the state at time $k+i$ as predicted at time $k$, and $N$ is the prediction horizon. The sets $\mathcal{X}\in\mathbb{R}^n$ and $\mathcal{U}\in\mathbb{R}^m$ define the constraints on the state and the input, respectively. Finally, the set $\mathcal{X}_f\subseteq\mathcal{X}$ defines the terminal constraint on the state. If we consider a regulation problem, the system (\ref{eqn:MPC_plant_discrete}) should be steered to the origin and the cost function $J(\vec{x}(k), \textbf{u})$ could be in a quadratic form as follows:
\begin{equation}
\label{eqn:MPC_cost_function_regulation}
	J(\vec{x}(k), \textbf{u}) = \vec{x}^\intercal_{k+N}\vec{P}_f\vec{x}_{k+N}+\sum_{i=1}^{N}\Big(\vec{x}^\intercal_{k+i}\vec{Q}\vec{x}_{k+i}+\vec{u}_{k+i}^\intercal\vec{R}\vec{u}_{k+i}\Big)
\end{equation}
where $\vec{P}_f,\vec{Q}\geq0$ (positive semi-definite) and $\vec{R}>0$ (positive definite) are weighting matrices.

@@

An MPC has many strengths. Given that the model is discrete and linear it handles multivariable problems very well. Also mathematical convexity is an important part of the resulting problem formulation of an MPC. In fact there exists efficient solvers for convex optimization problems but it is therefore desirable that the MPC problem (\ref{eqn:MPC_optimization}) is convex which is ensured if:
\begin{enumerate}
\item the cost function is convex;
\item the prediction model is linear;
\item the constraint sets $\mathcal{X}, \mathcal{U}$ are convex.	
\end{enumerate}

The optimization handles actuator constraints and state constraints naturally in the optimization which allows for
the process to be operated much closer to the hard constraints, which improves control performance and efficiency. Because of its predictive nature it is able to solve a variety of problems and handle disturbances smoothly.
